---
title: "Input Update Version 3"
author: "Chong Li"
date: "4/23/2022"
output: html_document
---


### FRS inputs Version 3

#### Document Purposes
* To make sure that all inputs included in microsimulation is included http://stage.nccp.org/tools/frs/test.php.
* To make sure that the inputs are coded into needed formats
  + https://docs.google.com/spreadsheets/d/170XpuxjgfXahEKFkuEzQ9V9mVSt8uW5A/edit#gid=1031769906
  + https://github.com/chong-cyber/nj_microsimulation/tree/main/NJmicrosimulation

#### Changes made
* `family_structure`: address the recoding mistake in version 2.
* `child[i]_age`: QUESTION - there are children whose age greater than 17, mostly in their 20s, what should be done with them?
* `parent[i]_age`: remove all families with ages greater than 62
  + The one entry with parent 1 less than 17
  + add the parent 3 and 4 ages, set them to be default 17.
* `married`: filtering for only married couples, add a variable call married
* `child_number`: create variable for number of children
* `family_size`: create variable for family size
* `disability_parent[i]` and `disability_child[i]`: add parent 3 and 4 disability to be 0.
* `cs_flag`: fixed previous errors, only child support for single parent families.
* `parent1_max_work`: numeric entries; for those who don't work, input 0.
* `maxshiftlength_parent1`: numeric entries; 
* `backtobackshifts_parent1`: set them to be integers between 0 to 2.
* `maxweekendshifts`: set to 1 for those who do have work on weekends
* `parent1_first_max`: set to the same as parent1_max_work
* `breadwinner_wkday_hometime`: set to the difference between parent 1 and 2 work hours.
* `breadwinner_wkend_hometime`: set to 0 for now; no real way to measure weekend hometime.
* `user_trans_type`: set the NAs to 0s.
* QUESTION: for the child support section, everything is set as FALSE or 0 and both.
* QUESTION: for benefits received if eligible, I have entered the benefits as the boxes are ticked on the NCCP website.
* `child care` section following the guidance on the excel sheet
* `privateplan_type` changes the NAs to "individual"
* Add `userplantype`
* QUESTION `fuel_source` seems to be either electric or gas in the source code, thus changed to these two.
* Set `override` to be 0s for a number of a different variables.
* Create `disability_work_expenses_m` variable.
* Create `wage_1` and `wage_parent2` as fraction of earnings variables.
* create `parent2_max_work_override_amt`; set it equal to `parent2_max_work`.
* Set `child_support_paid_m` to 0 for now.
* Set `heat_in_rent` as 0.
* Set all `itin` to 0 for now.
* 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,tidycensus,readxl,patchwork)
library(reshape2)

census_api_key("e0ff7586a0faf3bf1819178c1392ba6fa3df7ad9")
```

```{r}
# Read data
frs_individuals = read.csv("./data/cleaned_data/acs_individuals.csv")
frs_df = read.csv("./data/cleaned_data/frs_inputs_2.csv")
frs_df_old = read.csv("./data/cleaned_data/frs_inputs_1.csv")

pums_19var = pums_variables %>% 
  filter(survey == "acs1" & year== "2019") %>% 
  distinct(var_code, .keep_all = T) 
```

```{r}
nj_19h = haven::read_sas("./data/unix_hnj/psam_h34.sas7bdat")
```

```{r family structure}
# family_structure
frs_df_fs = frs_df %>%
  mutate(
    family_structure = frs_df_old$family_structure
  ) %>% 
  mutate(
    family_structure = ifelse(family_structure == "single-parent", 1, 2)
  )

```

```{r parent age }
# remove the parents who are older than 62 years old
old_parent = frs_df_fs %>%
  select(SERIALNO, ends_with("age")) %>% 
  filter(parent1_age > 62 | parent2_age > 62) %>% 
  distinct(SERIALNO, .keep_all = F)

frs_df_2 = frs_df_fs %>% 
  filter(!(SERIALNO %in% old_parent$SERIALNO)) %>% 
  mutate(
    parent3_age = 17, 
    parent4_age = 17
  )

```

```{r married}
# Only include households that are married
married_hh = nj_19h %>% 
  filter(HHT2 == "01" | HHT2 == "06" | HHT2 == "10")

frs_df_3 = frs_df_2 %>% 
  filter((SERIALNO %in% married_hh$SERIALNO)) %>% 
  mutate(
    married = ifelse(family_structure == 1, FALSE, TRUE)
  )
```

```{r child number}
child_num = frs_df_3 %>% 
  select(SERIALNO,child1_age:child5_age) %>% 
  pivot_longer(
    child1_age:child5_age,
    names_to = "child",
    values_to = "age"
  ) %>% 
  group_by(SERIALNO) %>% 
  count(kids_num = age != -1) %>% 
  filter(kids_num == T) 

frs_df_4 = frs_df_3 %>% 
  mutate(
    child_number = child_num$n
  )
```

```{r family size}
frs_df_5 = frs_df_4 %>% 
  mutate(
    family_size = child_number + family_structure
  )
```

```{r disability status}
frs_df_6 = frs_df_5 %>% 
  mutate(
    disability_parent3 = 0,
    disability_parent4 = 0
  )
```

```{r cs_flag}
cs_fl = frs_df_6 %>% 
  select(SERIALNO, family_structure, cs_flag) %>% 
  filter(
    family_structure == 1
  ) %>% 
  select(SERIALNO) %>% 
  cbind(cs_bin)

frs_df_7 = frs_df_6 %>% 
  left_join(cs_fl, by =  "SERIALNO")


frs_df_8 = frs_df_7 %>% 
  mutate(
    cs_flag = 0 
  ) %>% 
  mutate(
    cs_flag = ifelse(is.na(value), cs_flag, value)
  ) %>% 
  select(-value)
```

```{r parent1_max_work}
frs_df_9 = frs_df_8 %>% 
  mutate(
    parent1_max_work = ifelse(is.na(parent1_max_work), 0 , parent1_max_work)
  )

```

```{r maxshiftlength_parent1}
frs_df_10 = frs_df_9 %>% 
  mutate(
    maxshiftlength_parent1 = as.integer(parent1_max_work/5)
  ) 
```

```{r backtobackshifts_parent1}
frs_df_11 = frs_df_10 %>% 
  mutate(
    backtobackshifts_parent1 = ifelse(backtobackshifts_parent1 <= 0, 0, as.integer(backtobackshifts_parent1))
  )
```

```{r maxweekndshifts}
frs_df_12 = frs_df_11 %>% 
  mutate(
    maxweekendshifts = ifelse(weekenddaysworked != 0, 1, 0)
  )
```

```{r parent1_first_max}
frs_df_13 = frs_df_12 %>% 
  mutate(
    parent1_first_max = parent1_max_work
    ) 
```

```{r backtobackshifts_parent2}
frs_df_14 = frs_df_13 %>% 
  mutate(
    backtobackshifts_parent2 = ifelse(is.na(backtobackshifts_parent2),0,as.integer(backtobackshifts_parent2))
    )
```

```{r breadwinner_wkday_hometime}
frs_df_15 = frs_df_14 %>% 
  mutate(
    breadwinner_wkday_hometime = ifelse(family_structure == 2, abs(parent1_max_work - parent2_max_work), 0),
    breadwinner_wkend_hometime = 0
  ) 
```

```{r user_trans_type}
frs_df_16 = frs_df_15 %>% 
  mutate(
    user_trans_type = ifelse(is.na(user_trans_type),0, user_trans_type)
    )

frs_df_16 %>% count(user_trans_type)
```

```{r child support}
frs_df_17 = frs_df_16 %>% 
  mutate(
    ncparent_earnings = 0,
    cs_child_number = 0,
    child1support = FALSE,
    child2support = FALSE,
    child3support = FALSE,
    child4support = FALSE,
    child5support = FALSE,
    spousal_sup_cp = 0,
    spousal_sup_ncp = 0,
    child_support_ext_cp = 0,
    child_support_ext_ncp = 0
  )
```

```{r benefits received if eligible}
frs_df_18 = frs_df_17 %>% 
  mutate(
    ccdf = 1,fsp = 1,hlth = 1,sec8 = 0,tanf = 0,ctc = 1,eitc = 1,
    cadc = 1,ssi = 1,lifeline = 1,liheap = 1,wic = 1,prek = 1,ostp = 0,
    nsbp = 1,frpl = 1,fsmp = 1, premium_tax_credit = 1, ui = 0, headstart = 0, earlyheadstart = 0,
    state_cadc = 1, state_eitc = 1, prop_tax_credit = 1, state_mctr = 1, fli = 1, tdi = 1
    
  )
```

```{r child care settings}
# create child_care_nobenfit_estimate_source
frs_df_19 = frs_df_18 %>% 
  mutate(
    child_care_nobenefit_estimate_source = "spr",
    child_care_continue_estimate_source = "spr"
  ) 

# create child[i]_nobenefit_setting variables
c_prefix = paste0("child",1:5)
df_nobenefit = as.data.frame(matrix(nrow=8133,ncol=5)) 
colnames(df_nobenefit) = paste(c_prefix,"nobenefit_setting",sep="_") 
df_nobenefit = df_nobenefit %>% 
  replace(is.na(.),"accredited_center")

# create child[i]_nobenfit_amt_m variables
df_no_amt = as.data.frame(matrix(nrow=8133,ncol=5)) 
colnames(df_no_amt) = paste(c_prefix,"nobenfit_amt_m",sep="_") 
df_no_amt = df_no_amt %>% 
  replace(is.na(.),0)

# create child[i]_continue_setting
df_cont = df_nobenefit %>% 
  rename_at(1:5, funs(paste(c_prefix,"continue_setting",sep="_")))

# create child[i]_continue_amt_m variables
df_cont_amt = df_no_amt %>% 
  rename_at(1:5, funs(paste(c_prefix,"continue_amt_m",sep="_")))

# create child[i]_withbenefit_setting
df_bene = df_nobenefit %>% 
  rename_at(1:5, funs(paste(c_prefix,"withbenefit_setting",sep="_")))

# create child[i]_continue_flag
df_cont_fl = df_cont_amt %>% 
  rename_at(1:5, funs(paste(c_prefix,"continue_flag",sep="_")))

frs_df_19.1 = frs_df_19 %>% 
  cbind(df_nobenefit,df_no_amt,df_cont,df_cont_amt,df_bene,df_cont_fl)
```

```{r private_plan_type}
frs_df_20 = frs_df_19.1 %>% 
  mutate(
    privateplan_type = ifelse(is.na(privateplan_type), "individual", "employer"),
    userplantype = "employer"
    )
```

```{r fuel_source}
frs_df_21 = frs_df_20 %>% 
  mutate(
    fuel_source = case_when(
      fuel_source == 1 ~ "gas",
      fuel_source == 2 ~ "gas",
      fuel_source == 3 ~ "electricity"
    )
  ) %>% 
  mutate(
    fuel_source = ifelse(is.na(fuel_source), "other", fuel_source)
  ) 

```

```{r overrides}
frs_df_22 = frs_df_21 %>% 
  mutate(
    food_override = 0,
    trans_override = 0,
    other_override = 0
  )

```

```{r disability work expenses}
frs_df_23 = frs_df_22 %>% 
  mutate(
    disability_work_expenses_m = disability_personal_expenses_m/2
  )
```

```{r wage variables}
frs_df_24 = frs_df_23 %>% 
  mutate(
    wage_1 = earnings_initial * 0.6,
    wage_parent2 = earnings_initial * 0.4
    ) 
```

```{r parent2_max_work_override_amt}
frs_df_25 = frs_df_24 %>% 
  mutate(
    parent2_max_work_override_amt = parent2_max_work,
    child_support_paid_m = 0
  )
```

```{r additional variables to be set to 0}
frs_df_26 = frs_df_25 %>% 
  mutate(
    heat_in_rent = 0,
    headstart = 0,
    earlyheadstart = 0,
    prop_tax_credit = 0,
    food_override_amt = 0
  )
```

```{r maxworkweek_parent2}
frs_df_27 = frs_df_26 %>% 
  mutate(
    maxworkweek_parent2 = as.integer(parent2_max_work/8),
    maxworkweek_parent2 = ifelse(maxworkweek_parent2 > 7, 7, maxworkweek_parent2)
  ) 
```

```{r immigration status}
frs_df_28 =frs_df_27 %>% 
  mutate(
    across(child1_immigration_status:child5_immigration_status, 
           ~case_when(.x == 1 ~ "US Born",
                      .x == 2 ~ "US Territory Born",
                      .x == 3 ~ "Oversea Born US Citizen",
                      .x == 4 ~ "naturalized citizen",
                      .x == 5 ~"non-citizen",
                      .x == -9 ~ "0"))
  ) %>% 
  mutate(
    parent1_immigration_status = immigration,
    parent2_immigration_status = immigration,
    itin = 0
  ) 
```

```{r felony status}
frs_df_29 = frs_df_28 %>% 
  mutate(
    parent1_felony_conviction_status = NA
    ,parent2_felony_conviction_status = NA
    )
```

```{r out of pocket healthcare costs}
frs_df_30 = frs_df_29 %>% 
  mutate(
    fli = 0,
    tdi = 0,
    parent1_hlth_costs_oop_m = hlth_costs_oop_m,
    parent2_hlth_costs_oop_m = 0,
    child1_hlth_costs_oop_m = 0,
    child2_hlth_costs_oop_m = 0,
    child3_hlth_costs_oop_m = 0,
    child4_hlth_costs_oop_m = 0,
    child5_hlth_costs_oop_m = 0,
  ) 
```

```{r}
frs_df_30 %>% 
  mutate(
    other_override_amt = 0
  ) %>% 
  select(other_override_amt, snap_training, tanfwork, fullday_prek, fullday_k, recent_tanf_exit_flag)
```

```{r tanf receipients}
tanf_recp = frs_individuals %>% 
  select(SERIALNO, PAP) %>% 
  filter(PAP > 0) %>% 
  distinct(SERIALNO, .keep_all = T)

frs_df_31 = frs_df_30 %>% 
  left_join(tanf_recp, by = "SERIALNO") %>% 
  mutate(
    recent_tanf_exit_flag = ifelse(is.na(PAP), 0, 1)
  ) %>% 
  select(-PAP)
```

```{r}
write.csv(frs_df_31, "./data/cleaned_data/frs_inputs_3.csv", row.names = F)
```

```{r}
frs_df_31 %>% 
  count(workdaystart)
```



