---
title: "Input Update Version 6"
author: "Chong Li"
date: "5/27/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,tidycensus,readxl,patchwork)
library(reshape2)

census_api_key("e0ff7586a0faf3bf1819178c1392ba6fa3df7ad9")
```

```{r}
# Read data
frs_individuals = read.csv("./data/cleaned_data/acs_individuals.csv")
frs_df = read.csv("./data/cleaned_data/frs_inputs_5.csv")
frs_df_old = read.csv("./data/cleaned_data/frs_inputs_1.csv")

pums_19var = pums_variables %>% 
  filter(survey == "acs1" & year== "2019") %>% 
  distinct(var_code, .keep_all = T) 
```

```{r}
nj_19h = haven::read_sas("./data/unix_hnj/psam_h34.sas7bdat")
```

## add one column to the far left
```{r}
frs_df1 = frs_df %>% mutate(
  TESTCASEKEY = "s123456"
) %>% 
  select(TESTCASEKEY, everything())
```

## change fuel source for electricity
```{r}
frs_df2 = frs_df1 %>% 
  mutate(
    fuel_source = ifelse(fuel_source == "electricity", "electric", fuel_source)) 
```


## filter out those with parent1_age == 17

* There is one entry with such age.
```{r}
frs_df3 = frs_df2 %>% 
  filter(parent1_age != 17) 

frs_df3
```

## edit entries with parent2_age <= 17
* Those arel ikely households with married reference, but somehow the spousal information is missing
* Therefore adjusting them all to 1.
* also set the all age below 18 to 17.

```{r}
frs_df4 = frs_df3 %>% 
  mutate(
    family_structure = ifelse(family_structure == 2 & parent2_age <= 17, 1, family_structure)
  ) %>% 
  mutate(
    parent2_age = ifelse(parent2_age <= 17, 17, parent2_age)
  ) 
```
## issue with workhours_initial
* This is actually inferred from `WKHP`, which is a datapoint from ACS.
* Wage is actually calculated after earnings and WKHP.

```{r}
frs_df4 %>% 
  select(SERIALNO, parent1_earnings_initial,parent2_earnings_initial, parent1_workhours_initial, parent2_workhours_initial)
```

## max_shiftlength adjustments
* Now the reason why there are certain maxshiftlength at 0 is because the parent#_max_work is artificially set to 40. 
* I set the maxshiftlength to be 4.
```{r}
frs_df5 = frs_df4 %>% 
  mutate(maxshiftlength_parent1 = 
           ifelse(maxshiftlength_parent1 == 0 & parent1_max_work > 0, 4, maxshiftlength_parent1)
         ) %>% 
  mutate(maxshiftlength_parent2 = 
           ifelse(maxshiftlength_parent2 == 0 & parent2_max_work > 0, 4, maxshiftlength_parent2)
         )
```

## maxworkweek variable
* The issue also comes from the fact that many people work fewer than 40 hours per week - and the max_work was artificially set to 40 hours.
* set those for parent1 to 5, parent 2 to max_work/8
```{r}
frs_df6 = frs_df5 %>% 
  mutate(
    maxworkweek_parent1 = ifelse(maxworkweek_parent1 == 0 & parent1_max_work > 0, 5, maxworkweek_parent1)
  ) %>% 
  mutate(
    maxworkweek_parent2 = ifelse(maxworkweek_parent2 == 0 & parent2_max_work > 0, 1, maxworkweek_parent2)
  ) 
```

## tanf
* also set the savings to 1000
```{r}
tanf_fam = frs_individuals %>% 
  filter(PAP > 0) %>% 
  distinct(SERIALNO)

frs_df7 = frs_df6 %>% 
  mutate(
    tanf = ifelse(SERIALNO %in% tanf_fam$SERIALNO, 1, 0)
  ) %>% 
  mutate(
    savings = ifelse(tanf == 1, 1000, savings)
  )
```

## hlth
```{r}
pub_coverage = frs_individuals %>% 
  select(SERIALNO, PUBCOV) %>% 
  filter(PUBCOV == 1) %>% 
  distinct(SERIALNO)

frs_df8 = frs_df7 %>% 
  mutate(
    hlth = ifelse(SERIALNO %in% pub_coverage$SERIALNO, 1, 0)
  )
```

## FS

```{r}
food_stamp = nj_19h %>% 
  select(SERIALNO, FS) %>% 
  filter(FS == 1) %>% 
  distinct(SERIALNO)

frs_df9 = frs_df8 %>% 
    mutate(
    fsp = ifelse(SERIALNO %in% food_stamp$SERIALNO, 1, 0)
  )
```

## cep missing variables

```{r}
cep_data = read.csv("./Njmicrosimulation/FRS_Locations.csv")
```

```{r}
cep_par = cep_data %>% 
  select(name, cep_particpation) %>% 
  rename(
    residence_nj = name
  ) %>% 
  distinct(residence_nj, .keep_all = T)

frs_df10 = frs_df9 %>% 
  inner_join(cep_par, by = "residence_nj") %>% 
  rename(cep = cep_particpation)

```

## ethnicity variables
```{r}
single_parent_hh = nj_19h %>% 
  filter(HHT2 == "06" | HHT2 == "10" )

single_parent = frs_individuals %>%
  filter(SERIALNO %in% single_parent_hh$SERIALNO) %>%
  select(SERIALNO, AGEP, DIS, RELSHIPP, RAC1P) %>%
  filter(
    RELSHIPP == 20
  ) %>%
  arrange(SERIALNO,AGEP) %>%
  select(SERIALNO,AGEP,DIS,RAC1P) %>%
  mutate(
    id2 = ave(as.character(SERIALNO), SERIALNO, FUN = seq_along)
  )
```

```{r}
# select for the families with both parents around 
double_parent_hh = nj_19h %>% 
  filter(HHT2 == "01" | HHT2 == "03" )

# when you have both parents, parent 1 is the one with higher income
double_parent = frs_individuals %>%
  filter(SERIALNO %in% double_parent_hh$SERIALNO) %>%
  select(SERIALNO, DIS, AGEP, RELSHIPP,PINCP,SEX, RAC1P) %>%
  filter(
    RELSHIPP == 20 | RELSHIPP == 21 | RELSHIPP == 22 | RELSHIPP == 23 | RELSHIPP == 24
  ) %>% 
  group_by(SERIALNO) %>% 
  arrange(SERIALNO, desc(PINCP)) %>%
  select(SERIALNO,AGEP,PINCP,DIS,RAC1P) %>% 
  mutate(
    id2 = ave(as.character(SERIALNO), SERIALNO, FUN = seq_along)
  )

# convert the dataframe into one that indicate respective parent ID within each family.
parent_b4_trans = dcast(double_parent, id2 ~ SERIALNO, value.var="RAC1P") %>% 
  mutate(
    id2 = as.numeric(id2)
  ) %>% 
  arrange(id2) 

# transpose the dataframe for it to match with the original format
parent_a4 = as.data.frame(t(parent_b4_trans[,-1])) %>% 
    rename_at(1:2, funs(paste(prefix,"ethnicity",sep="_"))) %>% 
  mutate(
    SERIALNO = rownames(.)
  ) %>% 
  rename(
    parent1_ethnicity_1 = parent1_ethnicity,
    parent2_ethnicity_1 = parent2_ethnicity
  )
```

```{r}
prefix = paste0("parent",1:2)
columnsToAdd_1 = paste(prefix,"ethnicity",sep="_")

# single parent race
frs_df11 = frs_df10 %>% 
  add_column(!!!set_names(as.list(rep(NA, length(columnsToAdd_1))), nm=columnsToAdd_1)) %>% 
 left_join(single_parent, by = "SERIALNO") %>% 
  mutate(
    parent1_ethnicity = RAC1P,
    parent2_ethnicity = case_when((!is.na(RAC1P)) ~ NA)
  ) %>% 
  select(-c(AGEP,id2,DIS, RAC1P))

frs_df11.1 = frs_df11 %>% 
  left_join(parent_a4, by = "SERIALNO") %>%
  mutate(
    parent1_ethnicity = ifelse(is.na(parent1_ethnicity_1), parent1_ethnicity, parent1_ethnicity_1),
    parent2_ethnicity = ifelse(is.na(parent2_ethnicity_1), parent2_ethnicity, parent2_ethnicity_1)
  ) 

```

```{r}
# select RELSHIPP = 25, 26 and 27
new_ind = frs_individuals %>% select(SERIALNO, AGEP, DIS, RELSHIPP, CIT, RAC1P) %>% 
  filter(
    RELSHIPP == 25 | RELSHIPP == 26 | RELSHIPP == 27
  ) %>% 
  arrange(SERIALNO,AGEP) %>% 
  select(SERIALNO,AGEP,DIS,RELSHIPP,CIT, RAC1P) %>% 
  mutate(
    id2 = ave(as.character(SERIALNO), SERIALNO, FUN = seq_along)
  ) 

# extract the age of the kids, however, does not take account into families with 5+ kids (which exist)
# this is the non-transposed version
kids_b4_trans = dcast(new_ind, id2 ~ SERIALNO, value.var="RAC1P") %>% 
  mutate(
    id2 = as.numeric(id2)
  ) %>% 
  arrange(id2) %>% 
  filter(id2 <= 5) 

prefix = paste0("child",1:5)

# transpose the age dataframe, with family SERIALNO as the index
kids_a4 = as.data.frame(t(kids_b4_trans[,-1])) %>% 
    rename_at(1:5, funs(paste(prefix,"ethnicity",sep="_"))) %>% 
  mutate(
    SERIALNO = rownames(.)
  )

frs_df11.2 = frs_df11.1 %>% 
  left_join(kids_a4, by = "SERIALNO") 

frs_df11.2
```

```{r}
write.csv(frs_df11.2, "./data/cleaned_data/frs_inputs_6.csv", row.names = F)
```

