---
title: "Imputation SIPP"
author: "Chong Li"
date: "4/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse,tidycensus,readxl,patchwork)
library(reshape2)
library(mice)
require("data.table")
require("bit64")

census_api_key("e0ff7586a0faf3bf1819178c1392ba6fa3df7ad9")
```


### Document Purpose

* Document the steps of multiple imputation using SIPP data as (impute from survey), with ACS pUMS as (impute to survey)


#### STEP 1: Read in data
* FRS inputs
* ACS household/individual level from New Jersey
* SIPP selected data from New Jersey

```{r read data}
frs_individuals = read.csv("./data/cleaned_data/acs_individuals.csv")
frs_df = read.csv("./data/cleaned_data/frs_inputs_3.csv")
nj_19h = haven::read_sas("./data/unix_hnj/psam_h34.sas7bdat")
```

```{r read data from SIPP}
# 1. SSUID/PNUM/MONTHCODE and TEHC_ST 2. TAGE 4. EEDUC 5. EMS_EHC 6. ERELRPE 7. ERACE 8. ESEX 9. TMWKHRS 10. TPTOTINC 11. EBORNUS 12. TIMSTAT
require("data.table")
require("bit64")

pu = fread("./data/pu2020.csv", sep="|", select = c('SSUID','PNUM','MONTHCODE','TEHC_ST','TAGE','EEDUC',
                      'EMS_EHC', 'ERELRPE', 'ERACE', 'ESEX', 'TMWKHRS', 'TPTOTINC',
                      'EBORNUS',"ECITIZEN","ENATCIT",'TIMSTAT'))
```

#### STEP 2: Data Preprocessing

```{r data extraction}
# transform ACS individual data; remove extra variables
frs_individuals1 = frs_individuals %>% 
  select(-c(RT,DIVISION,SPORDER,REGION,ADJINC)) %>% 
  mutate(
    id = paste0(SERIALNO,"_",PUMA)
  ) %>% 
  select(id, everything()) %>% 
  select(-PUMA) %>% 
  select(-c(PWGTP1:PWGTP80))

# Select covariates and outcome (individual income) for imputation
# 1. person reference/state 2. age 3. class of work (unnecessary) 4. education level 5. marital status 6. relationship 7. race. 8. gender. 9. work hours 10. personal income 11. Nativity 12. 
acs_model = frs_individuals1 %>% 
  select(id, SERIALNO, AGEP, SCHL, MSP, RELSHIPP, RAC1P, SEX, WKHP, PINCP, NATIVITY, CIT) %>% 
  mutate(
    immigration_status = NA
  )

# Limit the SIPP data to only New Jersey entries; extract monthly data from October
# create new identifier, remove unnecessary variables
sipp_model = pu %>% 
  filter(TEHC_ST == 34) %>% 
  distinct(SSUID,PNUM,.keep_all=T) %>% 
  mutate(
    id = paste0(SSUID,"_", PNUM)
  ) %>% 
  select(-MONTHCODE, -TEHC_ST, -PNUM, -ENATCIT) %>% 
  relocate(id, everything())
```

```{r acs variable processing}
# rename and recode the ACS data
acs_model1 = acs_model %>% 
  rename(
    household_id = SERIALNO, age = AGEP, education = SCHL, marital_status = MSP, relationship = RELSHIPP,
    race = RAC1P, sex = SEX, work_hrs = WKHP, income = PINCP, nativity = NATIVITY, citizenship = CIT
  ) %>% 
  mutate(
    # recode education
    education = case_when(education == 1 ~ "no schooling",education == 2 ~ "preschool",
                          education == 3 ~ "kindergarten",education == 4 ~ "grade1",
                          education == 5 ~ "grade2", education == 6 ~ "grade3", 
                          education == 7 ~ "grade4", education == 8 ~ "grade5", education == 9 ~ "grade6",
                          education == 10 ~ "grade7" ,education == 11 ~ "grade8" , 
                          education == 12 ~ "grade9" , education == 13 ~ "grade10" , 
                          education == 14 ~ "grade11" , education == 15 ~ "grade12_no_diploma" ,
                          education == 16 ~"high school diploma", education == 17 ~ "GED" , 
                          education == 18 ~ "some college",education == 19 ~"1 or more year college, no degree",
                          education == 20 ~"associates' degree", education == 21 ~ "bachelor's degree",
                          education == 22 ~ "master's degree", education == 23 ~ "professional degree beyond bachelor's", education == 24 ~"doctorate degree", is.na(education) ~ "under 3 years old")) %>% 
  mutate(
    # recode marital status
    marital_status = case_when(
      marital_status == 1 ~ "married, spouse present", marital_status == 2 ~ "married, spouse absent",
      marital_status == 3 ~ "widowed", marital_status == 4 ~ "divorced", marital_status == 5 ~ "separated",
      marital_status == 6 ~ "nevermarried", is.na(marital_status) ~ "under 15 years old"
    )
  ) %>% 
  mutate(
    relationship = case_when(
      relationship == 20 ~ "reference", relationship == 21 ~ "opposite-sex spouse", 
      relationship == 22 ~ "opposite-sex partner", relationship == 23 ~ "same-sex spouse", 
      relationship == 24 ~ "same-sex partner", relationship == 25 ~ "child", relationship == 26 ~ "child", 
      relationship == 27 ~ "child", relationship == 28 ~ "sibling", relationship == 29 ~ "parent", relationship == 30 ~ "grand-child", relationship == 31 ~ "in-law-relatives", relationship == 32 ~ "in-law-relatives", 
      relationship == 33 ~ "other relative", relationship == 34 ~ "roommate", relationship == 35 ~ "child", 
      relationship == 36 ~ "other nonrelative", relationship == 37 ~ "roomer/border", relationship == 38 ~ "roomer/border"
    ) 
  ) %>% 
  mutate(
    race = case_when(
      race == 1 ~ "White", race == 2 ~ "Black", race == 3 ~ "Native American", race == 4 ~ "Alaaska Native",
      race == 5 ~ "Native American specified", race == 6 ~ "Asian", race == 7 ~ "Hawaiian Native",
      race == 8 ~ "Some other races", race == 9 ~ "two or more races"
    )
  ) %>% 
  mutate(
    sex = ifelse(sex == 1, "male", "female")
  ) %>% 
  mutate(
    nativity = ifelse(nativity == 1, "native", "foreigner")
  ) %>% 
  mutate(
    citizenship = case_when(
      citizenship == 1 ~ "U.S born citizen", citizenship == 2 ~ "U.S territory born citizen",
      citizenship == 3 ~ "born abroad to American parents", citizenship == 4 ~ "citizen by naturalization",
      citizenship == 5 ~ "not a citizen of U.S."
    )
  ) %>% 
  mutate(
    work_hrs = ifelse(age > 16 & is.na(work_hrs),0,work_hrs)
  )
  # mutate(
  #   education = as.factor(education),
  #   marital_status = as.factor(marital_status),
  #   relationship = as.factor(relationship),
  #   race = as.factor(race),
  #   sex = as.factor(sex),
  #   nativity = as.factor(nativity),
  #   citizenship = as.factor(citizenship))


```

```{r sipp variable processing}
# rename the variables in SIPP data so that it aligns with acs_model
colnames(sipp_model) = colnames(acs_model1)

sipp_model1 = sipp_model %>% 
  mutate(
    education = case_when(
      education == 31 ~ "Less than 1st grade", education == 32 ~ "1st, 2nd, 3rd or 4th grade", education == 33 ~ "5th or 6th grade", education == 34 ~ "7th or 8th grade",
      education == 35 ~ "9th grade", education == 36 ~ "10th grade", education == 37 ~ "11th grade", education == 38 ~  "12th grade, no diploma", 
      education == 39 ~"high school diploma/GED", education == 40 ~ "some college", education == 41 ~"1 or more year college, no degree",
      education == 42 ~"associates' degree", education == 43 ~ "bachelor's degree",
      education == 44 ~ "master's degree", education == 45 ~ "professional degree beyond bachelor's", education == 46 ~"doctorate degree",
    )
  ) %>% 
  mutate(
      marital_status = case_when(
          marital_status == 1 ~ "married, spouse present", marital_status == 2 ~ "married, spouse absent",
          marital_status == 3 ~ "widowed", marital_status == 4 ~ "divorced", marital_status == 5 ~ "separated",
          marital_status == 6 ~ "nevermarried", is.na(marital_status) ~ "under 15 years old"
    )
  ) %>% 
  mutate(
    relationship = case_when(
      relationship == 1 ~ "reference", relationship == 2 ~ "reference", relationship == 3 ~ "opposite-sex spouse", 
      relationship == 4 ~ "opposite-sex partner", relationship == 5 ~ "same-sex spouse", relationship == 6 ~ "same-sex partner", 
      relationship == 7 ~ "child", relationship == 8 ~ "grand-child", relationship == 10 ~ "sibling", 
      relationship == 9 ~ "parent", relationship == 11 ~ "in-law-relatives", relationship == 12 ~ "in-law-relatives",
      relationship == 13 ~ "other relative",relationship == 14 ~ "other relative", relationship == 16 ~ "roommate", relationship == 15 ~ "child", 
      relationship == 18 ~ "other nonrelative", relationship == 17 ~ "roomer/border")
  ) %>% 
  mutate(
    race = case_when(
      race == 1 ~ "white", race == 2 ~ "black", race == 3 ~ "asian", race == 4 ~ "other"
    )
  ) %>% 
  mutate(
    sex = ifelse(sex == 1, "male", "female")
  ) %>% 
  mutate(
    nativity  = ifelse(nativity == 1, "native", "foreigner")
  ) %>% 
  mutate(
    citizenship = ifelse(citizenship == 1, "citizen", "Non-citizen")
  ) %>% 
  mutate(
    immigration_status = ifelse(immigration_status == 1, "permanent", "other")
  ) %>% 
  mutate(
    education = ifelse(age < 3, "under 3 years old", education)
  ) %>% 
  mutate(
    income = 12*income
  ) %>% 
  mutate(
    work_hrs = ifelse(is.na(work_hrs) & age > 16, 0, work_hrs)
  ) %>% 
  mutate(
    income = ifelse(age < 15, NA, income)
  ) 
  # mutate(
  #   education = as.factor(education),
  #   marital_status = as.factor(marital_status),
  #   relationship = as.factor(relationship),
  #   race = as.factor(race),
  #   sex = as.factor(sex),
  #   nativity = as.factor(nativity),
  #   citizenship = as.factor(citizenship),
  #   immigration_status = as.factor(immigration_status)
  # )
```

```{r match ACS inputs with SIPP inputs}
acs_model2 = acs_model1 %>% 
  mutate(
    education = case_when(
      education == "under 3 years old" ~ "under 3 years old",
      education == "grade1" |education == "grade2"| education == "grade3"|education == "grade4" ~ "1st, 2nd, 3rd or 4th grade",
      education == "preschool" | education == "no schooling" | education == "kindergarten" ~ "Less than 1st grade",
      education == "grade5" | education == "grade6" ~ "5th or 6th grade",
      education == "grade7" | education == "grade8" ~ "7th or 8th grade",
      education == "grade9" ~ "9th grade",
      education == "grade10" ~ "10th grade",
      education == "grade11" ~ "11th grade",
      education == "grade12_no_diploma" ~ "12th grade, no diploma",
      education == "high school diploma" | education == "GED" ~ "high school diploma/GED",
      education == "some college" ~ "some college", education == "1 or more year college, no degree" ~ "1 or more year college, no degree",
     education == "associates' degree" ~ "associates' degree", education == "bachelor's degree" ~ "bachelor's degree",
     education == "master's degree" ~ "master's degree" , education == "professional degree beyond bachelor's" ~ "professional degree beyond bachelor's" , education == "doctorate degree" ~ "doctorate degree"
     )
  ) %>% 
  mutate(
    race = case_when(
      race == "White" ~ "white", race == "Black" ~ "black", race=="Asian" ~ "asian", 
      (race != "White" & race != "Black" & race != "Asian") ~ "other"
    )
  ) %>% 
  mutate(
    citizenship = ifelse(citizenship == "born abroad to American parents" | citizenship == "citizen by naturalization" |
                            citizenship == "U.S born citizen" | citizenship == "U.S territory born citizen","citizen","Non-citizen")
  ) 
```

**ACS**
* `id` is combo of SERIALNO and individual num
* `household_id` is the SERIALNO
* `education` has 17 unique level; complete
* `marital_status` has 7 values; complete
* `relationship` has 13 unique values; complete
* `race` has 4 unique values; complete. `sex` has 2 unique values; complete
* `nativity` and `citizenship` both have two values; complete
* `age` continuous and complete; `work_hrs` has complete rate of 0.578, incomplete values are for age < 16.
* `income`has complete rate of 0.629; incomplete values are for age < 15
* 

```{r ACS data exclusion/inclusion criterea}
acs_model2 %>% 
  skimr::skim()
```

**SIPP**

* * `id` is combo of USSID and individual num
* `household_id` is the USSID
* `education` has complete rate of 0.854; age< 15 are the NA values (COULD BE IMPUTED)
* `marital_status` is complete
* `relationship` is incomplete - due to being partial referenced/or temporarily outside of U.S./military (COULD BE IMPUTED)
* `race` has 4 unique values; complete. `sex` has 2 unique values; complete
* `nativity` and `citizenship` both have two values; complete
* `immigration_status`; 2 unique values; NA for non-natives
* `age` continuous and complete variable;
* `work_hrs` complete rate of 0.808, NA values are age < 16
* `income`has complete rate of 0.816; incomplete values are for age < 15

```{r sipp data exclusion/inclusion critiria}
sipp_model1 %>% 
  skimr::skim()
```

#### STEP 3: Within Survey Imputation

##### Model Specifications

**Is MAR assumption plausible**
* YES. Because we know that within the NA values only arise out of certain age/nativity exclusion criteria. Within the different strata, there are no missing values (hence MCAR).
* the reason for missing entries are not unknown to us.

**Form of imputation model**

**Variables to include as predictors**
* everything but the id

**impute variables that are functions of other variables**
* YES. given that they are relevant to the variables of interest

**variable imputation sequence**
* Check by the number of missing values

**starting imputations and # of iterations**
* need to monitor the convergence of MICE algorithm

**choose the number of multiply imputed datasets**
* 
```{r combine the two datasets}
# combine the ACS and the SIPP data-sets
combined_data = acs_model2  %>% 
  rbind(sipp_model1) %>% 
  mutate(
    education = as.factor(education),
    marital_status = as.factor(marital_status),
    relationship = as.factor(relationship),
    race = as.factor(race),
    sex = as.factor(sex),
    nativity = as.factor(nativity),
    citizenship = as.factor(citizenship),
    immigration_status = as.factor(immigration_status)
  )  %>% 
  mutate(
    data_source = as.factor(ifelse(str_detect(id,"[A-Z]"),"ACS","SIPP"))
  ) 
```


```{r prediction matrix}
# generate prediction matrix
pred_mat = quickpred(combined_data[3:ncol(combined_data)], mincor = 0.1)
```

```{r conditional imputation rules}
# set rules on conditional imputation
post = make.post(combined_data[3:ncol(combined_data)])

# for those younger than 3 years old, set education to three years old
post["education"] <- "imp[[j]][data$age[!r[, j]] < 3, i] <- levels(sipp_model2$education)[17]"

# for those who are natives,do not impute immigration status
post["immigration_status"] <- "imp[[j]][data$nativity[!r[, j]] == levels(data$nativity)[2], i] <- NA"

# for age < 16, no need for work hours
post["work_hrs"] <- "imp[[j]][data$age[!r[, j]] < 16, i] <- NA"

# for age < 15, no need for income
post["income"] <- "imp[[j]][data$age[!r[, j]] < 15, i] <- NA"
```

```{r conditional imputation using mice package}
combined_imp = mice((combined_data[3:ncol(combined_data)]), m = 1, predictorMatrix = pred_mat, 
                visitSquence = "monotone", maxit = 5, 
                post = post, seed =1111)
```

```{r extract complete imputation data}
# depending on the number of m's; the id cols might have to rbind() itself m times.

complete_imputation = combined_imp %>% 
  mice::complete() %>% 
  cbind(combined_data[1:2]) %>% 
  relocate(id,household_id,everything())
```

```{r}
xyplot(combined_imp,  immigration_status ~ data_source | .imp,
       )#|.imp,)
```

```{r}
sipp_model2 %>% 
  filter(citizenship == "citizen" & nativity == "foreigner" & is.na(immigration_status))
```


